---
title: "Answer 1 for Session1 part1: data summary"
author: "MRC Clinical Sciences Centre (http://mrccsc.github.io/)"
date: "14/Nov/2016"
output: html_document
---

These exercises cover the sections of Statistics in R [Session1: data summary]("../Session1_describe_the_data.html").

**Exercise 1**

- Read in the file Life_expectancy_at_birth.csv.
```{r,message=T}
 Life_exp<-read.csv("../data/Life_expectancy_at_birth.csv",header=T)
```
- Use the *head()* function to show the first 6 rows of the data. How many variables are there?
```{r,message=T}
 head(Life_exp)
```
- Use the *str()* function to see the data type of each column.
```{r,message=T}
 str(Life_exp)
```
- Use the *summary()* function to show the summarized information of the dataset.
```{r,message=T}
 summary(Life_exp)
```
- How many males and females are in the area of England.
```{r,message=T}
 Eng.life.exp<-Life_exp[Life_exp$area=="England",]
 ftable(Eng.life.exp$gender)
```
- Which function will you use to see the maximum and minumum of the life.expectancy in area England? 
```{r,message=T}
 # you can use min() and max() functions
 c(min(Eng.life.exp[,"life.expectancy"]),max(Eng.life.exp[,"life.expectancy"]))
 # or just use the range() function
 range(Eng.life.exp[,"life.expectancy"])
```
- What are the median life expenctancies of of Males and Females, respectively, in England?
```{r,message=T}
 median(Eng.life.exp[Eng.life.exp$gender=="Males","life.expectancy"])
 median(Eng.life.exp[Eng.life.exp$gender=="Females","life.expectancy"])
```
- What are the quantiles of life.expectancy in area Wales, N.Ireland, and in Scotland, respectively.
```{r,message=T}
 quantile(Life_exp[Life_exp$area == "Wales","life.expectancy"])
 quantile(Life_exp[Life_exp$area == "N.Ireland","life.expectancy"])
 quantile(Life_exp[Life_exp$area == "Scotland","life.expectancy"])
```
- show the variance and standard deviation of the life.expectancy in England
```{r,message=T}
 var(Life_exp[Life_exp$area == "England","life.expectancy"])
 sd(Life_exp[Life_exp$area == "England","life.expectancy"])
```
