---
title: "Answer for Session3: Correlation and Linear Regression"
author: "MRC Clinical Sciences Centre (http://mrccsc.github.io/)"
date: "14/Nov/2016"
output: html_document
---

These exercises cover the sections of Statistics in R [Session3: Correlation and Linear Regression]("../Session3_linear_regression.html").

**Exercise 1**

Use the *Petal.Width* (X) to predict the *Petal.Length* (Y) from the iris data

- load the *iris* data
```{r}
data(iris)
```
- What is the formula for the best fit lines for X vs Y

```{r,message=T}
 lmXY <- lm(Petal.Length~Petal.Width,iris)
lmXY
 plot(Petal.Length~Petal.Width,iris)
 abline(lmXY,col="red")

```

- What is the Rsquared for *Petal.Width* (X) vs *Petal.Length* (Y)
```{r,message=T}
summary(lmXY)
summary(lmXY)$r.squared
```

- Predict the *Petal.Length* when *Petal.Width=c(0.6,0.7,1.8)*
```{r,message=T}
new_width<-c(0.6,0.7,1.8)
# option 1: use the $coefficients
lmXY$coefficients
predictedLength <- unname(lmXY$coefficients[1] + lmXY$coefficients[2]*new_width)
predictedLength
# option 2: use the predict() function
dfnew_width<-data.frame(Petal.Width=c(0.6,0.7,1.8))
predict(lmXY,dfnew_width)
```

