---
title: "Answer for Session1 part2: data summary"
author: "MRC Clinical Sciences Centre (http://mrccsc.github.io/)"
date: "14/Nov/2016"
output: html_document
---

These exercises cover the sections of Statistics in R [Session1: data summary]("../Session1_describe_the_data.html").

**Exercise 2**

- Use the **pnorm()** function to see the probablity of a value being less than or equal to 2 for a normal distribution of mean 0 and standard deviation 1
```{r,message=T}
 pnorm(2,mean=0,sd=1)
```

- What is the probablity of a value being bigger than 2 for a normal distribution of mean 0 and standard deviation 1 (hint: use **?pnorm** in the Console and read the "help" tab in the lower right panel)
```{r,message=T}
# option 1
  pnorm(2,mean=0,sd=1,lower.tail = F)
# option 2
  1-pnorm(2,mean=0,sd=1)
```

**bonus question (optional)**

- load the gene expression matrix that has been created for this exercise from "data/gene_exp_matrix.RData"

```{r}
load("../data/gene_exp_matrix.RData")
head(gene_exp_matrix)
```

- use the gene expression to draw a heatmap
```{r}
#install.packages("gplots")
library("gplots")
heatmap.2(gene_exp_matrix,symkey=FALSE, trace="none",cexCol=0.75,cexRow=0.35,
          density.info="none",Colv = F,dendrogram="row")

```

- Use the *scale()* function to perform the Z-score transformation, and use the code above to generate the heatmap again (hint: read the help page of *?scale*, and you might need to use *t()* function as well)

```{r,message=T}
# option 1
scaled_gene_exp_matrix<-t(scale(t(gene_exp_matrix)))
heatmap.2(scaled_gene_exp_matrix,symkey=FALSE, trace="none",cexCol=0.75,cexRow=0.35,
          density.info="none",Colv = F,dendrogram="row")
```

However, the *heatmap.2()* function has a argument **scale**, which does the same thing for you....

